---
title: "linear model from Web"
author: "Kaylie Brehm"
date: "Summer 2022"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

 


```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

# Model 1 SIM~ARMS



### scatterplot

```{r}
plot(SIMS~ARM,data=data)
data1 = lm(SIMS~ARM,data=data)
abline(data1)
```

If you were to draw a straight line through the points, you would see that it is increasing and moving towards the right. So its a positive linear correlation correlation. The standard deviation is around 1 because it is a moderately tight scatterplot, with a relatively consistent bond width of about 4.

### Numerical results

```{r}
cor(SIMS~ARM,data=data)
```

The correlation is 0.6860073, which is fairly well coordinated.

### Inferential  (Build model.1)

```{r}
summary(data1)
newdata =c(SIMS=94, ARM=88)
```

The equation of the line is:   
SIMS = m * ARMS + b   
SIMS = 0.054563 * ARMS - 4.095160   

Our adjusted R-squared is 0.467, which indicates that ARM explains under half of the variation of the SIMS data. In other words, it reduces the error of the mean model by 46.7%. The residual standard error was calculated at 1.226, so the standard deviation is 1.226. This means that the bond width is 4.904, which is decently narrow. This indicates that ARMS can likely be used to predict SIMS.

#### Predict at target point

```{r}
predictSIM=predict.lm(data1, newdata=data.frame(SIMS=94, ARM=88))
print(predictSIM)
```

The prediction is about 0.71 for an ARM value of 88.

#### scatterplot with model fit
  
```{r}
plot(data1)
```
  
  
At any point above the line is a normal curve with a standard deviation of 1.226.

# Model 2 SIM~GRIP

### Now add in scatterplot

```{r}
plot(SIMS~GRIP,data=data)
data2 = lm(SIMS~GRIP,data=data)
abline(data2)
```


Just as in the former case, there seems to be an uphill trend to the right, indicative of a positive linear correlation. However, GRIP has a smaller correlation with SIMS then with SIMS vs ARM.

### Numerical results 

```{r}
cor(SIMS~GRIP,data=data)
```

Here the correlation of SIMS and GRIP is 0.6398 which is a bit smaller than the correlation of SIMS and ARM.

### Inferential  (Build model.2)


```{r}
model.2 <- lm(SIMS~GRIP,data=data)
summary.lm(model.2)
``` 

Our adjusted R-squared is 0.4053, which indicates that GRIP explains less than half of the variation of the SIMS data. It reduces the error of the mean model by 40.53%. The residual standard error was calculated at 1.295, which is equal to the standard deviation. This means the bond width is 1.295 * 4, or 5.18. Because the bond width is so narrow, it can be used to predict SIMS. But because it is greater than the previous example with ARM, it is less accurate in its predictions.



#### predict model.2 at target point

```{r}
predictSIM2=predict.lm(data2, newdata=data.frame(SIMS=94, GRIP=88))
print(predictSIM2)
```

#### now add the model fit to our plot for model.2
  
```{r}
plot(data2)
```

# Model 3 SIM~ARM+GRIP

### Numerical results (cor)

```{r}
data3=lm(SIMS~ARM+GRIP, data=data)
summary(data3)
```
  
### Inferential  (Build 2-dimentional model.3)

```{r}
model.2 <- lm(SIMS~ARM + GRIP,data=data)
summary.lm(model.2)
newdata =c(SIMS=94, ARM=88)
newdata =c(SIMS=94, GRIP=88)
``` 

#### predict model.3 at target point
  
```{r}
predict.lm(data3, newdata=data.frame(GRIP=94, ARM=88))
```
  

## Comparing nested models ANOVA Test

### Model.1 vs Model.3

```{r}
anova(data1, data3)
```


### Model.2 vs Model.3

```{r}
anova(data2, data3)
```

## Informally compare Model.1 with model.2

