---
title: "linear model from Web"
author: "Kaylie Brehm"
date: "Summer 2022"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

 


```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

# Model 1 SIM~ARMS



### scatterplot

```{r}
plot(SIMS~ARM,data=data)
data1 = lm(SIMS~ARM,data=data)
abline(data1)
```


### Numerical results

```{r}
cor(SIMS~ARM,data=data)
```

### Inferential  (Build model.1)

```{r}
summary(data1)
newdata =c(SIMS=94, ARM=88)
```

#### Predict at target point

```{r}
predictSIM=predict.lm(data1, newdata=data.frame(SIMS=94, ARM=88))
print(predictSIM)
```


#### scatterplot with model fit
  
```{r}
plot(data1)
```
  
  


# Model 2 SIM~GRIP

### Now add in scatterplot

```{r}
plot(SIMS~GRIP,data=data)
data2 = lm(SIMS~GRIP,data=data)
abline(data2)
```

### Numerical results 

```{r}
cor(SIMS~GRIP,data=data)
```

### Inferential  (Build model.2)

```{r}
summary(data2)
newdata =c(SIMS=94, GRIP=88)
```

#### predict model.2 at target point

```{r}
predictSIM2=predict.lm(data2, newdata=data.frame(SIMS=94, GRIP=88))
print(predictSIM2)
```

#### now add the model fit to our plot for model.2
  
```{r}
plot(data2)
```

# Model 3 SIM~ARM+GRIP

### Numerical results (cor)

```{r}
data3=lm(SIMS~ARM+GRIP, data=data)
summary(data3)
```
  
### Inferential  (Build 2-dimentional model.3)

```{r}
model.2 <- lm(SIMS~ARM + GRIP,data=data)
summary.lm(model.2)
newdata =c(SIMS=94, ARM=88)
newdata =c(SIMS=94, GRIP=88)
``` 

#### predict model.3 at target point
  
```{r}
predict.lm(data3, newdata=data.frame(GRIP=94, ARM=88))
```
  

## Comparing nested models ANOVA Test

### Model.1 vs Model.3

```{r}
anova(data1, data3)
```


### Model.2 vs Model.3

```{r}
anova(data2, data3)
```

## Informally compare Model.1 with model.2

