---
title: "linear model from Web"
author: "Kaylie Brehm"
date: "Summer 2022"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

 


```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

# Model 1 SIM~ARMS



### scatterplot

```{r}
plot(SIMS~ARM,data=data)
data1 = lm(SIMS~ARM,data=data)
abline(data1)
```

If you were to draw a straight line through the points, you would see that it is increasing and moving towards the right. So its a positive linear correlation correlation. The standard deviation is around 1 because it is a moderately tight scatterplot, with a relatively consistent bond width of about 4.

### Numerical results

```{r}
cor(SIMS~ARM,data=data)
```

The correlation is 0.6860073, which is fairly well coordinated.

### Inferential  (Build model.1)

```{r}
summary(data1)
newdata =c(SIMS=94, ARM=88)
```

The equation of the line is:   
SIMS = m * ARMS + b   
SIMS = 0.054563 * ARMS - 4.095160   

Our adjusted R-squared is 0.467, which indicates that ARM explains under half of the variation of the SIMS data. In other words, it reduces the error of the mean model by 46.7%. The residual standard error was calculated at 1.226, so the standard deviation is 1.226. This means that the bond width is 4.904, which is decently narrow. This indicates that ARMS can likely be used to predict SIMS.

#### Predict at target point

```{r}
predictSIM=predict.lm(data1, newdata=data.frame(SIMS=94, ARM=88))
print(predictSIM)
```

The prediction is about 0.71 for an ARM value of 88.

#### scatterplot with model fit
  
```{r}
plot(data1)
```
  
  
At any point above the line is a normal curve with a standard deviation of 1.226.

# Model 2 SIM~GRIP

### Now add in scatterplot

```{r}
plot(SIMS~GRIP,data=data)
data2 = lm(SIMS~GRIP,data=data)
abline(data2)
```


Just as in the former case, there seems to be an uphill trend to the right, indicative of a positive linear correlation. However, GRIP has a smaller correlation with SIMS then with SIMS vs ARM.

### Numerical results 

```{r}
cor(SIMS~GRIP,data=data)
```

Here the correlation of SIMS and GRIP is 0.6398 which is a bit smaller than the correlation of SIMS and ARM.

### Inferential  (Build model.2)


```{r}
model.2 <- lm(SIMS~GRIP,data=data)
summary.lm(model.2)
``` 

Our adjusted R-squared is 0.4053, which indicates that GRIP explains less than half of the variation of the SIMS data. It reduces the error of the mean model by 40.53%. The residual standard error was calculated at 1.295, which is equal to the standard deviation. This means the bond width is 1.295 * 4, or 5.18. Because the bond width is so narrow, it can be used to predict SIMS. But because it is greater than the previous example with ARM, it is less accurate in its predictions.



#### predict model.2 at target point

```{r}
predictSIM2=predict.lm(data2, newdata=data.frame(SIMS=94, GRIP=88))
print(predictSIM2)
```

he prediction is about -0.81 for a GRIP value of 88.

#### now add the model fit to our plot for model.2
  
```{r}
plot(data2)
```

At any point above the line is a normal curve with a standard deviation of 1.295. 

# Model 3 SIM~ARM+GRIP

### Numerical results (cor)

```{r}
data3=lm(SIMS~ARM+GRIP, data=data)
summary(data3)
```
  
The correlation between SIMS~ARM was 0.6860073. The correlation between SIMS~GRIP was  0.6398458. If these two correlations were considered fairly large, then the correlation between SIMS and ARM+GRIP is even greater, standing at 0.7326545. This is a notably large correlation
  
### Inferential  (Build 2-dimentional model.3)

```{r}
model.2 <- lm(SIMS~ARM + GRIP,data=data)
summary.lm(model.2)
newdata =c(SIMS=94, ARM=88)
newdata =c(SIMS=94, GRIP=88)
``` 

SIMS = .037311(ARM) + .024470(GRIP)-5.433871

If we look at the adjusted R squared value of this GRIP data, it is smaller than the ARM data adjusted R squared value. In addition, the residual standard error is larger for this grip data then the ARM data. Both of these values indicate that model 1 is a better model.

#### predict model.3 at target point
  
```{r}
predict.lm(data3, newdata=data.frame(GRIP=94, ARM=88))
```
  
  When ARM is 88 and GRIP is 94, the prediction of SIMS is 0.149. The 95% confidence interval goes from -2.132373 to 2.431668--a distance of 4.564041. This is 2 standard deviations up and 2 standard deviations down. Since the residual standard error is 1.144, which is approximately one fourth of the 95% confidence interval.

## Comparing nested models ANOVA Test

### Model.1 vs Model.3

```{r}
anova(data1, data3)
```

The output shows that in model 1 the amount of errors in the mean model is 217.88 but when we include ARM + GRIP, the error in the mean model decreased to 188.43. So, the model with two regressions in it is superior to the model with just one.

### Model.2 vs Model.3

```{r}
anova(data2, data3)
```

Here model 3 is shown to be better than model 2 as it shows that the error in the GRIP model is 243.07.

## Informally compare Model.1 with model.2

While no P-values are available, since they are nested models, you can still informally compare them. The amount of error between Model 1 and 2 is 25.19. This indicates that ARM better explains SIMS than GRIP, since it has a noticeably lesser amount of error. Model 1 is a better model than model 2 as evidence of RSS values of 243.07 vs 188.43. 
